\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{xcolor}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!50!black,
  urlcolor=blue!50!black,
  citecolor=blue!50!black
}

\setlist[itemize]{leftmargin=1.25em}
\setlist[enumerate]{leftmargin=1.5em}

\title{\textbf{Methodology Whitepaper}\\React Native Evals Benchmark}
\author{Repository-aligned methodology specification}
\date{February 26, 2026}

\begin{document}
\maketitle

\begin{abstract}
This document specifies the methodology implemented by the React Native Evals repository for benchmarking coding models on React Native tasks. It describes dataset structure, execution flow, solver and judge roles, requirement-weighted scoring, aggregation logic, artifact outputs, and reproducibility considerations. The description is implementation-faithful to the current repository state and explicitly notes branch-level authoring conventions where they affect interpretation.
\end{abstract}

\section{Scope and Assumptions}
This whitepaper documents the benchmark as executed through the split CLI flow: \texttt{bun runner/run.ts} followed by \texttt{bun runner/judge.ts}. It is aligned to observable runner behavior under \texttt{runner/} and to repository-level eval authoring guidance under \texttt{docs/}.

\textbf{Assumption used in this document:} evals follow the repository contract under \texttt{evals/<category>/<eval-id>/} and are benchmarked via the standard runner pipeline.

\section{Benchmark Unit and Dataset Contract}
The atomic benchmark unit is a single eval directory. Each eval is expected to contain:
\begin{itemize}
  \item \texttt{prompt.md}: task prompt used by the solver model
  \item \texttt{requirements.yaml}: a list of judgeable requirements (optional per-requirement weights)
  \item \texttt{app/}: baseline input files provided to the solver
  \item \texttt{reference/}: reference implementation files retained for authoring and manual validation workflows
\end{itemize}

Eval discovery is file-driven: the runner scans for \texttt{requirements.yaml}. If that file is missing, the eval is not discovered and will not be included in the run.

\section{End-to-End Execution Pipeline}
For each benchmark run, execution is split into generation and judging:
\begin{enumerate}
  \item \textbf{Generation CLI}: discover evals, load \texttt{app/} and \texttt{prompt.md}, run solver model, write generated files and \texttt{manifest.json} to the configured output directory.
  \item \textbf{Judge CLI}: read eval entries from generation \texttt{manifest.json}, load generated files and \texttt{requirements.yaml}, run LLM judging, write per-eval results incrementally under \texttt{runs/<input-folder>/evals/}, and write summary at run completion.
\end{enumerate}

The runner supports bounded concurrency. If an eval fails and \texttt{----fail-fast} is not enabled, the run continues and the eval is counted in \texttt{evalsErrored}.

\section{Solver Stage Methodology}
\subsection{Input Construction}
The solver is given:
\begin{itemize}
  \item the eval prompt from \texttt{prompt.md}
  \item the baseline files from \texttt{app/}
  \item a fixed system instruction requiring the model to return all provided files with modifications applied
\end{itemize}

The runner formats baseline files into a structured prompt with explicit file tags.

\subsection{Output Contract}
The solver returns structured output with:
\begin{itemize}
  \item \texttt{summary}: short description of performed changes
  \item \texttt{files[]}: array of \texttt{\{path, content\}}
\end{itemize}

Returned paths are sanitized before writing to disk (e.g., stripping leading slashes and removing traversal segments such as \texttt{..}).

\subsection{Model Requirement}
The generation CLI requires \texttt{----model}. When set to \texttt{noop}, the generation stage copies files from each eval's \texttt{reference/} directory into the configured output and still writes a validated \texttt{manifest.json}. Any other model value runs normal solver generation.

\section{LLM Judge Methodology}
\subsection{Requirements Parsing}
\texttt{requirements.yaml} is parsed and validated at runtime. The enforced schema requires:
\begin{itemize}
  \item \texttt{version}: positive integer (defaults to 1 via schema default)
  \item \texttt{requirements[]}: non-empty list of objects with:
  \begin{itemize}
    \item \texttt{id} (non-empty string)
    \item \texttt{description} (non-empty string)
    \item optional positive \texttt{weight}
  \end{itemize}
\end{itemize}

Extra fields are ignored by the current runtime parser.

\subsection{Deterministic \texttt{implementation-} Requirement Convention}
The repository documents an authoring convention for deterministic requirements:
\begin{itemize}
  \item prefix deterministic, source-verifiable requirement IDs with \texttt{implementation-}
  \item use \texttt{weight: 1} by default to avoid overweighting implementation details
\end{itemize}

Examples include required imports, API usage patterns, and specific wiring that can be verified directly from files.

\textbf{Important distinction:} the runner does \emph{not} infer behavior from ID prefixes such as \texttt{implementation-}. Prefixes are an authoring convention; scoring changes only when a requirement explicitly sets \texttt{weight} in \texttt{requirements.yaml}.

\subsection{Technical Eval Authoring Controls}
For technical categories (for example navigation, animation, storage), requirement quality is governed by additional authoring controls documented under \texttt{docs/}.
\begin{itemize}
  \item \textbf{Recent API shift audit:} before requirement writing, review official docs plus recent releases/changelogs for each core library, and extract concrete ``new preferred API'' and ``deprecated/removed API'' signals with dates.
  \item \textbf{Evidence-gated \texttt{MUST NOT}:} use explicit \texttt{MUST NOT} clauses only when backed by primary-source deprecation/removal or correctness caveats.
  \item \textbf{Implementation-level specificity:} technical diversity is measured at requirement/API-constraint level, not prompt wording level.
  \item \textbf{Similarity budget:} within a library subgroup, keep shared baseline requirements small (target: at most two shared IDs) and maintain multiple eval-specific implementation constraints per eval.
  \item \textbf{Judge input scope:} set \texttt{inputs.files} to implementation files under evaluation (for example \texttt{app/App.tsx}) rather than prompt-only artifacts.
\end{itemize}

These controls are authoring policy for dataset quality; they are not independently enforced by runner schema validation.

\subsection{Judge Prompt Construction}
The judge prompt includes:
\begin{itemize}
  \item declared requirements (including weights when present)
  \item generated files under evaluation
  \item rules instructing the judge to use only provided files as evidence and return one result per declared requirement ID
\end{itemize}

\subsection{Structured Judge Output}
The judge returns structured output with:
\begin{itemize}
  \item optional \texttt{summary}
  \item \texttt{requirements[]}, where each row includes \texttt{id}, \texttt{passed}, \texttt{reason}, \texttt{evidence[]}, and optional \texttt{confidence}
\end{itemize}

\subsection{Requirement Mapping and Failure Policy}
Judge rows are mapped back to declared requirements by ID.
\begin{itemize}
  \item If a declared requirement is missing from judge output, the runner marks it as failed.
  \item The runner emits a fixed failure reason indicating that the judge did not return a result for that requirement.
  \item Extra judge rows for undeclared IDs are ignored for scoring.
  \item Requirement weights are normalized before scoring (Section~\ref{sec:scoring}).
\end{itemize}

The judge CLI requires \texttt{----model}; there is no judge skip/noop path.

\section{Scoring Methodology}
\label{sec:scoring}
\subsection{Requirement Weight Normalization}
Each requirement weight is normalized as:
\[
w_i' =
\begin{cases}
w_i & \text{if } w_i \text{ is finite and } w_i > 0 \\
1 & \text{otherwise}
\end{cases}
\]

This normalization is defensive. In normal operation, the runtime schema already requires positive weights when provided.

\subsection{Per-Eval Weighted Requirement Score}
For an eval with declared requirements \(R\), normalized weights \(w_i'\), and binary pass indicators \(p_i\):
\[
\text{totalWeight} = \sum_{i \in R} w_i'
\]
\[
\text{passedWeight} = \sum_{i \in R} p_i \cdot w_i'
\]
\[
\text{scoreRatio} =
\begin{cases}
0 & \text{if } \text{totalWeight} = 0 \\
\frac{\text{passedWeight}}{\text{totalWeight}} & \text{otherwise}
\end{cases}
\]

The runner rounds \texttt{passedWeight}, \texttt{totalWeight}, and \texttt{scoreRatio} to 4 decimal places.

\textbf{Interpretation example.} If requirements have weights \([2, 2, 1]\) and pass outcomes \([1, 0, 1]\), then the per-eval score is:
\[
\frac{2 + 0 + 1}{2 + 2 + 1} = \frac{3}{5} = 0.6
\]
This pattern is relevant only when authors intentionally assign non-default weights to specific requirements.

\subsection{Run-Level Aggregate Score}
Let \(E^+\) be the set of evals that complete successfully (i.e., without an unrecovered pipeline error). The summary field \texttt{weightedAverageScore} is:
\[
\texttt{weightedAverageScore} =
\begin{cases}
0 & \text{if } |E^+| = 0 \\
\frac{1}{|E^+|}\sum_{e \in E^+}\texttt{scoreRatio}_e & \text{otherwise}
\end{cases}
\]
rounded to 4 decimal places.

\textbf{Interpretation note:} the ``weighted'' part refers to weights within each eval. The run-level average itself is an unweighted mean across successful evals.

\subsection{Additional Summary Metrics}
The run summary also reports:
\begin{itemize}
  \item \texttt{requirementsTotal}
  \item \texttt{requirementsPassed}
  \item \texttt{evalsErrored}
\end{itemize}

For successful evals, each row includes \texttt{requirementsTotal}, \texttt{requirementsPassed},
and \texttt{scoreRatio}. For pipeline errors, each row includes \texttt{status=error}
along with the eval identifier and path.

\section{Artifacts and Output Schema (Operational View)}
Each run writes artifacts to deterministic generation and judge directories:
\begin{itemize}
  \item \texttt{runs/<input-folder>/evals/<eval-id>.json}: per-eval judge results (written as each eval completes)
  \item \texttt{runs/<input-folder>/summary.json}: aggregate judge summary (written at end)
  \item generated artifacts and \texttt{manifest.json}: written by generation CLI to a user-selected output directory (default: \texttt{generated/<model>-<run-id>})
  \item per-eval OpenCode session snapshots:
  \begin{itemize}
    \item generation: \texttt{<generated-eval-dir>/opencode-session.solver.json}
    \item judging: \texttt{runs/<input-folder>/evals/<eval-id>.opencode-session.judge.json}
  \end{itemize}
  \item optional debug artifacts when \texttt{----debug} is enabled
\end{itemize}

Per-eval results include solver and judge model identifiers, mapped requirement verdicts, per-eval score, generated file list, and optional \texttt{judgeSessionArtifactPath} when session capture succeeds. The generation \texttt{manifest.json} includes optional \texttt{solverSessionArtifactPath} per eval.

\section{Recommended Reporting Protocol}
For comparative studies, report at minimum:
\begin{enumerate}
  \item repository commit hash (dataset and runner version)
  \item CLI options (\texttt{run: ----pattern, ----model, ----timeout, ----concurrency, ----output}; \texttt{judge: ----model, ----timeout, ----concurrency, ----input})
  \item execution date and time
  \item counts of discovered, processed, and errored evals
  \item \texttt{weightedAverageScore} and \texttt{requirementsPassed/Total}
  \item generated artifact source directory
\end{enumerate}

Repeat runs are recommended. The current runner does not explicitly pin a random seed or decoding settings (e.g., temperature) in solver or judge calls, so provider defaults and model nondeterminism can affect results.

\section{Limitations and Threats to Validity}
\begin{itemize}
  \item \textbf{Judge dependence:} requirement verdicts depend on the configured LLM judge model and prompt behavior.
  \item \textbf{Requirement-interpretation ambiguity:} without seeded reference context, ambiguous requirement phrasing can increase variance in judge decisions.
  \item \textbf{Binary requirement outcomes:} there is no partial credit within a single requirement.
  \item \textbf{Run-level averaging choice:} the summary score is unweighted across evals, which can underweight larger evals.
  \item \textbf{Pipeline error exclusion from mean:} errored evals are counted separately and excluded from the run-level score average.
  \item \textbf{Authoring-convention variability:} conventions such as \texttt{implementation-} improve consistency but are not enforced by the runtime parser.
\end{itemize}

\section{Conclusion}
The methodology combines structured solver outputs and requirement-based LLM judging to evaluate React Native coding performance on repository-defined tasks. Its main strength is transparency: scoring is based on declared requirements with explicit weights, documented mapping behavior, and reproducible artifact outputs. The documented \texttt{implementation-} convention provides clearer guidance for deterministic requirement authoring without changing the underlying runner logic.

\end{document}
